\documentclass{beamer}
\usetheme[hideothersubsections]{HRTheme}
\usepackage{beamerthemeHRTheme}
\usepackage{graphicx}
\usepackage[space]{grffile}
\usepackage{listings}
\lstset{language=SQL,
basicstyle=\ttfamily\footnotesize,
mathescape=true,
keywordstyle=\color{blue},
breaklines=true}
\usepackage[utf8]{inputenc}
\usepackage{color}
\newcommand{\red}[1]{
\textcolor{red}{#1}
}
\newcommand{\ts}{\textbackslash}

\title{E-R model, Relational model, SQL}

\author{ }

\institute{Hogeschool Rotterdam \\ 
Rotterdam, Netherlands}

\date{}

\begin{document}
\maketitle

\SlideSection{Introduction}
\SlideSubSection{Lecture topics}
\begin{slide}{
\item E-R model.
\item Relational model.
\item SQL, and examples.
}\end{slide}

\SlideSection{Data models}
\SlideSubSection{database design process}
\begin{slide}{
\item Requirements analysis 
\pause
\item Conceptual database design using E-R models
\pause
\item Logical database design (sometimes conceptual and logical are merged into one step) 
\pause
\item Schema refinement through normalization
\pause
\item Physical database design
\pause
\item Application and Security Design
}\end{slide}

\SlideSection{E-R model}
\SlideSubSection{Overview}
\begin{slide}{
\item Highest level of database modelling.
\item Model the conceptual aspect of the database.
\item Far from the physical representation in the DBMS.
}\end{slide}

\SlideSubSection{Entity}
\begin{slide}{
\item Anything which can exist on its own on the database
\item Consider a database for a space shooter game
\item Starships, asteroids are entities, they have a meaning on their own

%insert picture here
}\end{slide}

\SlideSubSection{Attributes}
\begin{slide}{
\item They model characteristics of the entity.
\item \textbf{Starship:} velocity, shield, armour, weapon, [...]
\item \textbf{Asteroid:} velocity, mass, integrity, [...]

%insert picture here
}\end{slide}

\SlideSubSection{Relations}
\begin{slide}{
\item They describe the associations among entities (two or more).
\item They have a cardinality: number of participants for each side.

%insert picture here
}\end{slide}

\SlideSubSection{Relations - 1 : 1}
\begin{slide}{
\item Entity modelling a pilot and one modelling a starship.
\item Related by ``drives''.
\item The cardinality is 1:1 : one pilot drives at most one starship, and one starship can contain only one pilot.

%insert picture here
}\end{slide}

\SlideSubSection{Relations - 1 : N}
\begin{slide}{
\item Entity modelling a starship and one modelling a weapon.
\item Realted by ``mounted''
\item The cardinality is 1:N : a weapon can be mounted only on one starship, but a starship can mount more than one weapon.

%insert picture here
}\end{slide}

\SlideSubSection{Relations - N : M}
\begin{slide}{
\item Entity modelling a starship and one modelling an asteroid.
\item Realted by ``collides with''
\item The cardinality is N : M : several starships can collide with several asteroids.

%insert picture here
}\end{slide}

\SlideSubSection{Keys}
\begin{slide}{
\item A way to uniquely identify an entity.
\item A key is a set of attributes that have unique values among entities.
\item \textbf{Starship:} id number.

%insert picture here
}\end{slide}

\SlideSubSection{Weak entities}
\begin{slide}{
\item Entities which do not have a key attribute.
\item \textbf{Asteroids:} There can be two asteroids with the same position, same mass, velocity, etc.

%insert picture here
}\end{slide}

\SlideSection{Relational model}
\SlideSubSection{Overview}
\begin{slide}{
\item Halfway between a conceptual model and the physical model.
\item Contain an abstraction of physical elements.
\item Can be easily mapped to a physical implementation in a DBMS.
\item There are mapping rules from E-R model to the relational model.
}\end{slide}

\SlideSubSection{Relation}
\begin{slide}{
\item A relation is a collection of tuples.
\item Each element of a tuple is a value taken from an attribute set.
\item Each attribute set is identified by a name

\vspace{0.5cm}
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{\textbf{Ships}} \\
\hline
\underline{id} & name & shields & armour & integrity \\
\hline
\end{tabular}

\vspace{0.5cm}
(38258269, "Battlestar Galactica", 3000, 5000, 1.0)
}\end{slide}

\SlideSubSection{Keys}
\begin{slide}{
\item A \textit{Primary key} is a set of attributes with unique values in each tuple.
\item A \textit{Candidate key} is the smallest set of attributes which form a superkey.

\vspace{0.5cm}
\textbf{Example:}\\
\textbf{Primary key:} (id, name, shield) \\
\textbf{Candidate key:} (id)
}\end{slide}

\SlideSubSection{Keys}
\begin{slide}{
\item A \textit{Primary key} is the chosen key for a relation among all the candidate keys.
\item A \textit{Foreign key} is a set of attributes in one relation which is a primary key in another relation.

\vspace{0.5cm}
\textbf{Example (Foreign key):}\\

\centering
\begin{tabular}{|c|c|}
\hline
\multicolumn{2}{|c|}{\textbf{Mounts}} \\
\hline
\underline{shipid} & \underline{weaponName} \\
\hline
\end{tabular}

\vspace{0.2cm}
\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{\textbf{Ships}} \\
\hline
\underline{id} & name & shields & armour & integrity \\
\hline
\end{tabular}

\raggedright
\vspace{0.2cm}
In the relation \texttt{Mounts} the attribute \texttt{shipid} is a foreign key to \texttt{Ships}.


}\end{slide}

\SlideSection{SQL}
\SlideSubSection{Overview}
\begin{slide}{
\item Used to create relations (tables).
\item Used to insert/modify/extract data from relations (tables).
\item Declarative language (``What'' not ``How'').
}\end{slide}

\SlideSubSection{SELECT-FROM}
\begin{frame}[fragile]
\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{\textbf{Ships}} \\
\hline
\underline{id} & name & shields & armour & integrity \\
\hline
\end{tabular}

\vspace{0.25cm}
\textbf{Select all ships from the game}
\pause
\begin{lstlisting}
SELECT *
FROM Ships
\end{lstlisting}

\end{frame}

\SlideSubSection{SELECT-FROM-WHERE (Selection)}
\begin{frame}[fragile]
\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{\textbf{Ships}} \\
\hline
\underline{id} & name & shields & armour & integrity \\
\hline
\end{tabular}

\vspace{0.25cm}
\textbf{Select all ships in the game whose pilot is ``William Adama''}
\pause
\begin{lstlisting}
SELECT *
FROM Ships s
WHERE s.pilot = 'William Adama'
\end{lstlisting}

\end{frame}

\SlideSubSection{SELECT-FROM-WHERE (Projection)}
\begin{frame}[fragile]
\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{\textbf{Ships}} \\
\hline
\underline{id} & name & shields & armour & integrity \\
\hline
\end{tabular}

\vspace{0.25cm}
\textbf{Find the name of the ships whose pilot is ``Starbucks''}
\pause
\begin{lstlisting}
SELECT s.name
FROM Ship s
where s.pilot = 'Starbucks'
\end{lstlisting}

\end{frame}

\SlideSubSection{JOIN}
\begin{frame}[fragile]

\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{\textbf{Ships}} \\
\hline
\underline{id} & name & shields & armour & integrity \\
\hline
\end{tabular}

\begin{tabular}{|c|c|}
\hline
\multicolumn{2}{|c|}{\textbf{Mounts}} \\
\hline
\underline{shipid} & \underline{weaponName} \\
\hline
\end{tabular}

\begin{tabular}{|c|c|c|}
\hline
\multicolumn{3}{|c|}{\textbf{Weapons}} \\
\hline
\underline{name} & damage & type \\
\hline
\end{tabular}

\vspace{0.25cm}
\textbf{Find the id of the ships mounting the weapon ``Stealthblade MKII''}
\pause
\begin{lstlisting}[showstringspaces=false]
SELECT s.id
FROM Ship s, Mounts m
WHERE s.id = m.shipid AND 
      m.weaponName = 'StealthBlade MKII'
\end{lstlisting}

\end{frame}

\SlideSubSection{JOIN (multiple tables)}
\begin{frame}[fragile]

\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{\textbf{Ships}} \\
\hline
\underline{id} & name & shields & armour & integrity \\
\hline
\end{tabular}

\begin{tabular}{|c|c|}
\hline
\multicolumn{2}{|c|}{\textbf{Mounts}} \\
\hline
\underline{shipid} & \underline{weaponName} \\
\hline
\end{tabular}

\begin{tabular}{|c|c|c|}
\hline
\multicolumn{3}{|c|}{\textbf{Weapons}} \\
\hline
\underline{name} & damage & type \\
\hline
\end{tabular}

\vspace{0.25cm}
\textbf{Find the name of all the weapons mounted in the ships flown by ``Apollo''}

\pause
\begin{lstlisting}[showstringspaces=false]
SELECT w.name
FROM Ship s, Mounts m, Weapon w
WHERE s.id = m.Shipid AND 
      m.weaponName = w.name AND
      s.pilot = 'Apollo'
\end{lstlisting}

\end{frame}

\SlideSubSection{Aggregates (SUM)}
\begin{frame}[fragile]

\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{\textbf{Ships}} \\
\hline
\underline{id} & name & shields & armour & integrity \\
\hline
\end{tabular}

\begin{tabular}{|c|c|}
\hline
\multicolumn{2}{|c|}{\textbf{Mounts}} \\
\hline
\underline{shipid} & \underline{weaponName} \\
\hline
\end{tabular}

\begin{tabular}{|c|c|c|}
\hline
\multicolumn{3}{|c|}{\textbf{Weapons}} \\
\hline
\underline{name} & damage & type \\
\hline
\end{tabular}

\vspace{0.25cm}
\textbf{Find the total damage output of the ships flown by ``Athena''}

\pause
\begin{lstlisting}[showstringspaces=false]
SELECT SUM(w.damage) AS damage
FROM Ship s, Mounts m, Weapon w
WHERE s.id = m.shipid AND 
      m.weaponName = w.name AND
      s.pilot = 'Athena'
\end{lstlisting}

\end{frame}

\SlideSubSection{Aggregates (COUNT, HAVING)}
\begin{frame}[fragile]

\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{\textbf{Ships}} \\
\hline
\underline{id} & name & shields & armour & integrity \\
\hline
\end{tabular}

\begin{tabular}{|c|c|}
\hline
\multicolumn{2}{|c|}{\textbf{Mounts}} \\
\hline
\underline{shipid} & \underline{weaponName} \\
\hline
\end{tabular}

\begin{tabular}{|c|c|c|}
\hline
\multicolumn{3}{|c|}{\textbf{Weapon}} \\
\hline
\underline{name} & damage & type \\
\hline
\end{tabular}

\vspace{0.25cm}
\textbf{Count all the ships having more than 3 weapons}

\pause
\begin{lstlisting}[showstringspaces=false]
SELECT COUNT(*)
FROM (
 SELECT COUNT(*) AS ShipCount
 FROM Ship s, Mounts m, Weapon w
 WHERE s.id = m.shipid AND 
      m.weaponName = w.Name
 GROUP BY s.id
 HAVING COUNT(*) > 3)
                   
\end{lstlisting}

\end{frame}






\end{document}

\begin{slide}{
\item ...
}\end{slide}

\begin{frame}[fragile]
\begin{lstlisting}
...
\end{lstlisting}
\end{frame}
