\section{Practicum assignments}
	In this section the practicum assignments and their grading criteria are listed.
	
	\paragraph*{Assignment 1}
		The student is asked to implement a simple database for a given real-world scenario. This database is managed by a management application written in a chosen OO-language (Java or C\#). The student is also ask to provide analysis on possible functional dependencies and provide a normalized version of the database.

	\paragraph*{Assignment 2}
		Implement the scenario of the first assignment in a document-based database. The implementation must come with a management application for the database using the map-reduce paradigm to execute given queries and to populate the database.

	\paragraph*{Assignment 3}
		Given the scenario described in the previous assignments, the student is asked to realise a correspondent graph database model, and implement it in graph DBMS Neo4j. The database must be populated by using the Cypher language coming with Neo4j, as well as the queries.
		
		\subsubsection*{Assignment grading:}
		The grading system is showed in the table below. The teacher reserves the right to assign half of the points if the goal is partially achieved. If the student is individually unable to answer a question about any of the assignments, the student will not receive the score for that assignment.
		
		Note that in Assignment 1 you can receive an extra point for the ERD but the total score cannot exceed the maximum amount of 4 points. Also note that the front end application connected to a database is a minimum requirement for the assignment to be graded. If that is missing, the grade will be automatically 0.
		
		
		\begin{longtable}{|p{3cm}|p{8cm}|}
			\hline
			\textbf{Assignment 1} &
			\begin{itemize}
				\item All the data is correctly modelled by the database (1 pt.).
				\item The report contains significant functional dependencies and an explanation about the normal form of each table, possibly showing normalization steps (1 pt.).
				\item All the operations listed in the assignment description are correctly implemented (1 pt.).
				\item The application uses design patterns to abstract the database layer from the application layer (1 pt.)
				\item There is a correct ERD modelling the database (extra 1 pt.) 
			\end{itemize} \\
			\hline
			\textbf{Assignment 2} &
			\begin{itemize}
				\item There is a random data generator (1 pt.)
				\item The map-reduce queries are correctly implemented 1 pts.)
				\item The student makes use of design patterns to change the database layer and recycle the code for the front end application (1 pt.)
			\end{itemize} \\
			\hline
			\textbf{Assignment 3a} &
			\begin{itemize}
				\item The graph database structure is correctly implemented (1 pt.)
				\item The data respects the given constraints (1 pt.)
				\item The queries in cypher are correctly implemented (1 pts.)
			\end{itemize} \\
			\hline
			\textbf{Assignment 3b} &
			\begin{itemize}
				\item The graph database structure is correctly implemented (1 pt.)
				\item The queries are correctly implemented (1 pt.)
				\item Enough entities and relationships have been mapped into the database (1 pt.)
			\end{itemize}\\
			\hline
		\end{longtable}
